---
title: "STAT 2200: Problem Set 5"
author: "Jack Ambery"
date: "Due: Wednesday, 3/27 at the beginning of class"
fontsize: 12pt
output: pdf_document
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


\vspace{0.1in}

\begin{itemize}
  \item You may discuss this assignment with other students in the class, but 
    you may not sit down and type it up with them or show them your code. You 
    also may not discuss the assignment with anyone who isn't in our class, nor
    may you look up anything online.
  \item You must type up your homework using R Markdown. I want to see all of 
    your code and output, and any answers you provide that aren't code must be 
    typed above the respective \texttt{R} code chunk.
  \item Make sure none of your code runs off the page, otherwise you will lose 
    points.
  \item You must print and turn in a PDF of your homework. I won't accept 
    anything else (e.g., a Word document). All pages must be stapled together.
  \item This assignment is worth 75 points.
\end{itemize}


\vspace{0.2in}


\noindent 1. Write a function that converts temperatures measured in degrees 
Celsius into temperatures measured in degrees Fahrenheit. The equation for doing 
so is: $F = \frac{9}{5}C+32$. Then use the function to convert the following 
three temperatures from degrees Celsius to degrees Fahrenheit: 0, 14, and 31. 
Print the resulting temperatures (in degrees Fahrenheit).
```{r}

```

\vspace{0.2in}

\noindent 2. The formula for calculating the amount of money someone has after
investing an initial amount of money in an account that earns interest 
compounded a certain number of times per year is: $A = P(1+\frac{r}{n})^{nt}$, 
where $A=$ the final amount, $P=$ the principal balance (the amount invested 
initially), $r=$ the annual nominal interest rate (as a proportion), $n=$ the 
number of times the interest is compounded per year, and $t=$ the number of 
years the entire amount is invested without any money being withdrawn from (or 
more added by the investor to) the account. Write a function that calculates $A$ 
using the formula above. Then use the function to calculate (and print) the 
amount of money someone has in their account after investing \$4,500 with a 
3.5\% nominal interest rate, compounded monthly, for 20 years.
```{r}

```

\vspace{0.2in}

\noindent 3. Write a function that accepts a vector of values and returns the 
following statistics as a named vector: mean, median, standard deviation, IQR, 
range, $2.5^{\text{th}}$ percentile, and $97.5^{\text{th}}$ percentile. Then use 
the function to find this set of statistics for the annual amount of 
precipitation in US cities as recorded in the \textit{precip} dataset, which 
comes with base \texttt{R}. Make sure the set of statistics is printed.
```{r}

```

\newpage

\noindent 4. Write a function with three arguments: \texttt{x}, \texttt{y}, and 
\texttt{z}, all of which are meant to be individual numbers. The function must 
do the following: 
\begin{itemize}
\item if the sum of \texttt{x} and \texttt{y} is under 10, then the function 
  returns the product of \texttt{y} and \texttt{z}, minus 2;
\item if the sum of \texttt{x} and \texttt{y} is between 10 and 15, inclusive, 
  then the function returns the sum of \texttt{x} and \texttt{z}, plus 6;
\item otherwise, the function returns the product of \texttt{x}, \texttt{y}, and 
  \texttt{z}.
\end{itemize}

Then use the function for the following three scenarios: (1) x = 1, y = 6, 
z = 3, (2) x = 8, y = 6, z = 2, and (3) x = 10, y = 7, z = 9. Make sure the 
three results are printed separately.
```{r}

```

\vspace{0.2in}

\noindent 5. Quadratic functions follow the form $f(x)=ax^2+bx+c$, and the graph 
of a quadratic function is a parabola. To find where a parabola passes through 
the $x$-axis (values known as roots), you set the equation equal to 0 and find 
the $x$ value(s) that satisfy the equation. There are three possibilities for 
the number of roots of a quadratic function: 0, 1, or 2. The quadratic formula, 
as you may remember from a previous math course, finds the root(s) of a 
quadratic function, if any exist. The formula is:
\begin{center}
$\displaystyle x=\frac{-b \pm \sqrt{b^2-4ac}}{2a}$.
\end{center}

It turns out that if the part under the square root (called the discriminant) is 
a negative number, then there are no roots; if the discriminant equals 0, then 
there is only one root; and if the discriminant is a positive number, then there 
are two roots. Note that I'm ignoring imaginary numbers here, for anyone who 
knows what those are.

Write a function -- with default values of 1 for $a$, 2 for $b$, and 1 for $c$ 
-- that does the following:
\begin{itemize}
\item if no root exists, then a message is printed saying there are no roots;
\item if only one root exists, then a message is printed saying there is only one 
  root, and the root must be printed as well; and
\item if two roots exist, then a message is printed saying there are two roots, 
  and both roots must be printed as well.
\end{itemize}

Then find the roots for the following quadratic functions: (1) 
$f(x)=5x^2-8.5x+2.8$, (2) $f(x)=-3x^2+5x-11$, and (3) $f(x)=4x^2-7x+12.25$. Make 
sure the result is printed for each of the three functions.
```{r}

```



