---
title: "STAT 2200: Problem Set 3"
author: "Jack Ambery"
date: "Due: Monday, 2/26 at the beginning of class"
fontsize: 12pt
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- set global options -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\vspace{0.1in}

```{=tex}
\begin{itemize}
  \item You may discuss this assignment with other students in the class, but 
    you may not sit down and type it up with them or show them your code. You 
    also may not discuss the assignment with anyone who isn't in our class, nor
    may you look up anything online.
  \item You must type up your homework using R Markdown. I want to see all of 
    your code and output, and any answers you provide that aren't code must be 
    typed above the respective \texttt{R} code chunk.
  \item Make sure none of your code runs off the page, otherwise you will lose 
    points.
  \item You must print and turn in a PDF of your homework. I won't accept 
    anything else (e.g., a Word document). All pages must be stapled together.
  \item This assignment is worth 100 points.
\end{itemize}
```
\vspace{0.1in}

In problems 1 and 2 you will work with datasets from the website for the
Lock, Lock, Lock, Lock, and Lock textbook
\textit{Statistics: Unlocking the Power of
Data} (3rd ed.). Go to the following URL to access the two datasets:
\url{https://www.lock5stat.com/datapage3e.html}. \vspace{0.1in}

\noindent 1. Download the CSV file that contains the \textit{Cars2020}
dataset and read it into \texttt{R}. Use \texttt{str()} to examine the
structure of the data set. Then:
```{r}
Cars2020 <- read.csv("../datasets/Cars2020.csv")
str(Cars2020)
```


a)  Make four plots -- one for each of the following variables: CityMPG,
    Drive, LowPrice, and Type. Make sure you're using an appropriate
    type of plot for each. Note: please don't make a pie chart for any
    categorical variable. You can earn a bonus point if you add
    appropriate labels to each plot.
```{r}
# CityMPG
hist(Cars2020$CityMPG, xlab = "CityMPG of Cars", 
     main = "Frequency of City MPGs")

# Drive
drives <- table(Cars2020$Drive)
barplot(drives, xlab = "Type of Drive", 
        main = "Frequency of Different Types of Drive")

# LowPrice
hist(Cars2020$LowPrice, xlab = "Low Prices of Cars", 
     main = "Frequency of Low Prices")

# Type
types <- table(Cars2020$Type)
barplot(types, xlab = "Type of Car", 
        main = "Frequency of Different Car Types")
```
    

b)  For the CityMPG variable, calculate the mean, median, standard
    deviation, IQR, range, $5^{\text{th}}$ percentile, and
    $95^{\text{th}}$ percentile. Then type these values outside of the
    \texttt{R} code chunk.
```{r}
mpgs <- Cars2020$CityMPG
summary(mpgs)
sd(mpgs)
iqr <- 19.0 - 13.25
iqr
range <- 28 - 10
range
quantile(mpgs, 0.05)
quantile(mpgs, 0.95)
```
__For the CityMPG variable__

Mean: 16.19

Median: 15.5

Standard Deviation: 3.74

IQR: 5.75

Range: 18

$5^{\text{th}}$ percentile: 11

$95^{\text{th}}$ percentile: 22.55

c)  For the Type variable, calculate (i) the frequency of cars that fall
    into each category, and (ii) the percentage of cars that fall into
    each category. Then type these values outside of the \texttt{R} code
    chunk.
```{r}
anyNA(Cars2020$Type) #FALSE
table(Cars2020$Type)
round(((table(Cars2020$Type)/nrow(Cars2020)) * 100), 2) #percentages
```
__Frequencies of Car Types__

Hatchback:

* Frequency = 4

* Percentage = 3.64%

Minivan:

* Frequency = 4

* Percentage = 3.64%

Sedan:

* Frequency = 38

* Percentage = 34.55%

Sporty:

* Frequency = 11

* Percentage = 10%

SUV:

* Frequency = 50

* Percentage = 45.45%

Wagon:

* Frequency = 3

* Percentage = 2.73%


\vspace{0.1in}

\noindent 2. Download the CSV file that contains the
\textit{Hurricanes2018} dataset and read it into \texttt{R}. Run
\texttt{str()} to examine the structure of the dataset. Next, make a
timeplot that plots the number of hurricanes over the past century or
so. Make sure the timeplot has points at each year and lines connecting
the points. You can earn a bonus point if you add a title and
appropriate labels to both axes.
```{r}
Hurricanes2018 <- read.csv("../datasets/Hurricanes2018.csv")
str(Hurricanes2018)
```

```{r}
plot(Hurricanes2018$Year, Hurricanes2018$Hurricanes, 
     xlab = "Year", 
     ylab = "Number of Hurricanes", 
     main = "Number of Hurricanes per Year")
lines(Hurricanes2018$Year, Hurricanes2018$Hurricanes)
```


\vspace{0.1in}

\noindent 3. Download the \textit{HollywoodMovies} dataset from
Blackboard. As you look at the dataset, you should notice missing
values. Read the file into \texttt{R} and run the \texttt{unique()}
function on the variable Genre. This function outputs all of the
unique/different entries that exist in an object. Do any entries stand
out as possibly having been read improperly? If so, reimport your data
properly. Next, use an appropriate type of plot to plot the domestic
revenue of the movies (DomesticGross) based on genre (Genre), but only
considering the following three genres: action, drama, and romance.
```{r}
HollywoodMovies <- read.csv("../datasets/HollywoodMovies.csv", 
                            na.strings = "")
```

```{r}
unique(HollywoodMovies$Genre)

genre <- HollywoodMovies$Genre
dgross <- HollywoodMovies$DomesticGross

boxplot(dgross[genre == "Action"], 
        dgross[genre == "Drama"], 
        dgross[genre == "Romance"], 
        ylab = "Domestic Gross", 
        xlab = "Movie Genres", 
        main = "Domestic Gross of Action, Drama, and Romance Movies", 
        names = c("Action", "Drama", "Romance"))

```


\vspace{0.1in}

\noindent 4. Download the \textit{mlb2011} dataset from Blackboard. This
file was originally found on BaseballGuru.com (located at
\url{https://BaseballGuru.com}) and contains data for Major League
Baseball (MLB) pitchers in 2011. In this problem you are going to look
at the relationship between the number of wins (variable "W") and the
number of strikeouts (variable "SOA") for pitchers with at least 10
games started (variable "GS") across the two different leagues (variable
"LG"). The two leagues are the National League (NL) and the American
League (AL).
```{r}
mlb2011 <- read.csv("../datasets/mlb2011.csv")
```


After reading the data into \texttt{R}, you first want to separate
pitchers into two groups: those who had at least 10 starts and played in
the National League, and those who had at least 10 starts and played in
the American League. Then create a scatterplot with the number of wins
on the $y$-axis and the number of strikeouts on the $x$-axis. You should
appropriately distinguish between the National League and American
League pitchers using points with different colors. Next, add two
regression lines (aka lines of best fit) to the plot. The first line,
which has a $y$-intercept of 0.945 and a slope of 0.073, corresponds to
American League pitchers. The second line, which has a $y$-intercept of
1.966 and a slope of 0.059, corresponds to National League pitchers.
Color the lines appropriately. You can earn a bonus point if you
correctly add axis labels and a legend to your plot.
```{r}
league <- mlb2011$LG
gamesStarted <- mlb2011$GS
NLPitchers <- mlb2011[league == 'NL' & gamesStarted >= 10,]
ALPitchers <- mlb2011[league == 'AL' & gamesStarted >= 10,]

plot(NLPitchers$SOA, NLPitchers$W, 
     col = 'red',
     main = 'Number of Strikeouts vs. Wins for 
     Pitchers in the MLB National and American Leagues in 2011',
     xlab = 'Number of Strikeouts',
     ylab = 'Number of Wins')
points(ALPitchers$SOA, ALPitchers$W, col='blue')
abline(a = 1.966, b = 0.059, col='red')
abline(a = 0.945, b = 0.073, col='blue') 
```

