---
title: "STAT 2200: Problem Set 4"
author: "Jack Ambery"
date: "Due: Monday, 3/18 at the beginning of class"
fontsize: 12pt
output: pdf_document
---


<!-- set global options -->
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




\begin{itemize}
  \item You may discuss this assignment with other students in the class, but 
    you may not sit down and type it up with them or show them your code. You 
    also may not discuss the assignment with anyone who isn't in our class, nor
    may you look up anything online.
  \item You must type up your homework using R Markdown. I want to see all of 
    your code and output, and any answers you provide that aren't code must be 
    typed above the respective \texttt{R} code chunk.
  \item Make sure none of your code runs off the page, otherwise you will lose 
    points.
  \item You must print and turn in a PDF of your homework. I won't accept 
    anything else (e.g., a Word document). All pages must be stapled together.
  \item This assignment is worth 100 points.
\end{itemize}


\vspace{0.1in}

\noindent 1. Create a sequence of 20 numbers as follows: the first number is 5, 
and each number thereafter is found by multiplying the current number by 1.5 and 
then subtracting 1. Print the resulting sequence.
```{r}
arr <- c(5, rep(0, times = 19))
for (i in 1:19)
  arr[i+1] <- (arr[i] * 1.5) - 1

print(arr)
```
\vspace{0.1in}

\noindent 2. Suppose you have the following vector of probabilities, each of 
which measures the probability a person has a particular disease: 0.25, 0.83, 
0.76, 0.25, 0.33, 0.51, 0.67, 0.50, 0.54, 0.75, 0.71, 0.03, 0.38, 0.64, and 
0.19. Create a new variable where a one indicates that the person is predicted 
to have the disease (based on a probability of 0.5 or above) and a zero 
indicates that the person is predicted not to have the disease (based on a 
probability under 0.5). Do this two different ways: 
\begin{itemize}
\item first, using a loop, and
\item second, using the \texttt{ifelse()} function.
\end{itemize}

Print the resulting vectors. Note that they should contain the exact same 
values!
```{r}
probabilities <- c(0.25, 0.83, 0.76, 0.25, 0.33, 0.51, 0.67, 0.50, 0.54, 0.75, 0.71, 0.03, 0.38, 0.64, 0.19)
likelyhoods <- vector()

for (i in 1:length(probabilities))
  if (probabilities[i] >= 0.5) {
    likelyhoods[i] <- 1
  } else {
    likelyhoods[i] <- 0
  }
print('Likelyhood using loop:')
print(likelyhoods)

likelyhoods2 <- ifelse(probabilities >= 0.5, 1, 0)
print('Likelyhood using ifelse:')
print(likelyhoods2)
```

\vspace{0.1in}

\noindent 3. In this problem you will work with the \textit{Cars2020} dataset,
which you downloaded from \url{http://www.lock5stat.com/datapage3e.html} and 
used on Homework 3. Please work with the CSV again here. Create a new variable 
that classifies the fuel efficiencies of the cars in the dataset based on the 
following made-up criteria: a highway mpg under 20 is considered poor, a highway 
mpg between 20 and 25 (inclusive) is considered acceptable, a highway mpg above 
25 and at most 35 is considered good, and a highway mpg above 35 is considered 
great. Make sure the dataset contains the new variable, and then print only the 
following for every car in the dataset: the make and model of the car, as well 
as the classified fuel efficiency.
```{r}
Cars2020 <- read.csv("../datasets/Cars2020.csv")

fuelEff <- Cars2020$HwyMPG
fuelClass <- vector()

for (i in 1:length(fuelEff)) {
  if (fuelEff[i] < 20) {
    fuelClass[i] = "poor"
  } else if (fuelEff[i] >= 20 & fuelEff[i] <= 25) {
    fuelClass[i] = "acceptable"
  } else if (fuelEff[i] > 25 & fuelEff[i] <= 35) {
    fuelClass[i] = "good"
  } else { #fuelEff[i] > 35
    fuelClass[i] = "great"
  }
}
# adding it to Cars2020 dataframe
Cars2020$fuelClass <- fuelClass

#printing out specific columns of Cars2020
print(Cars2020[c("Make","Model","fuelClass")])

```

\vspace{0.1in}

\noindent 4. Use a loop to fill a $6 \times 8$ (i.e., six rows and eight 
columns) matrix as follows: if the sum of the indices of an element's position 
in the matrix is even, then the new element is the product of the indices. On 
the other hand, if the sum of the indices is odd, then the new element is the 
sum of the indices plus three. Print the resulting matrix.
```{r}
table <- matrix(0, nrow = 6, ncol = 8)

for(i in 1:nrow(table)) {
  for (j in 1:ncol(table)) {
    if ((i + j) %% 2 == 0) { # if even
      table[i, j] <- i*j
    } else { # if odd
      table[i, j] <- i + j + 3
    }
  }
}
print(table)
```

\vspace{0.1in}
  
\noindent 5. Suppose I taught a course last semester that consisted of homework, 
two midterm exams, and a final exam, which are weighted as follows when 
calculating a student's final course average: homework is worth 35%, the 
midterms are each worth 20%, and the final exam is worth 25%. Also suppose I set 
up the course grading scheme as follows when determining final course letter 
grades:
\begin{itemize}
\item A = [90, 100]
\item B = [80, 90)
\item C = [70, 80)
\item D = [60, 70)
\item F = below 60
\end{itemize}

The grades for the students in the class can be found in the 
\textit{CourseGrades} dataset on Blackboard. Determine the letter grade each 
student earned in the course. That letter grade should be added as a new 
variable to the data frame, and you must print the final data frame. Note that 
everything in this problem is made up.
```{r}
grades <- read.csv("../datasets/CourseGrades.csv")
FinalGrade <- vector()
LetterGrade <- vector()

for (i in 1:nrow(grades)) {
  hwPercent <- grades[i, 2] * 0.35
  midterm1 <- grades[i, 3] * 0.2
  midterm2 <- grades[i, 4] * 0.2
  final <- grades[i, 5] * 0.25
  total <- hwPercent + midterm1 + midterm2 + final
  FinalGrade[i] <- total
}

for (i in 1:length(FinalGrade)) {
  if (FinalGrade[i] >= 90) {
    LetterGrade[i] = "A"
  } else if (FinalGrade[i] < 90 & FinalGrade[i] >= 80) {
    LetterGrade[i] = "B"
  } else if (FinalGrade[i] < 80 & FinalGrade[i] >= 70) {
    LetterGrade[i] = "C"
  } else if (FinalGrade[i] < 70 & FinalGrade[i] >= 60) {
    LetterGrade[i] = "D"
  } else { #FinalGrade[i] < 60
    LetterGrade[i] = "F"
  }
}

grades$LetterGrade <- LetterGrade
print(grades)
```

\vspace{0.1in}

\noindent 6. Suppose a friend of yours is currently a second-semester senior, 
and her GPAs for the previous (individual) semesters, in order (starting first 
semester freshman year), are as follows: 4.00, 3.50, 3.67, 3.33, 2.67, 3.42, and 
3.56. Assuming she took the same number of credits each semester, write a loop 
that calculates and stores her cumulative GPA after each semester. Then print 
the resulting vector of cumulative GPAs.
```{r}
gpas <- c(4.00, 3.50, 3.67, 3.33, 2.67, 3.42, 3.56)
cumulatives <- vector()

for (i in 1:length(gpas)) {
  cumulatives[i] <- mean(gpas[1:i])
}
print("Cumulative GPAs:")
cumulatives
```

\vspace{0.1in}

\noindent 7. Returning to problem 6, suppose your friend was on an academic 
scholarship that requires the student to maintain a cumulative GPA of at least 
3.6 in order to continue to receive financial support the following semester. If 
at any point the student's cumulative GPA drops below 3.6, they lose the 
scholarship and stop receiving the financial support permanently. Write a while 
loop that counts the number of semesters your friend received financial support 
from the scholarship during her college career. Print that number. Additionally, 
you should assume your friend received the benefits starting first semester 
freshman year.
```{r}
semesters = 1
cumulative = cumulatives[1]
while (cumulative >= 3.6) {
  cumulative = cumulatives[semesters + 2]
  semesters = semesters + 1
}
print("The student held the scholarship for this many semesters:")
print(semesters)

```

\vspace{0.1in}

\noindent 8. The Fibonacci sequence is a very special sequence that consists of
the following numbers: 0, 1, 1, 2, 3, 5, 8, 13, 21, 34, etc. If you aren't 
familiar with this sequence and don't know how each number is found, please 
search on Google. It's pretty cool! Write code that calculates the first 40
numbers in the sequence, and then print them. Note that you may start with the 
first two numbers in the sequence only.
```{r}
fib <- c(0, 1)
for (i in 3:40) {
  fib[i] <- fib[i - 1] + fib[i - 2]
}
print("First 40 of Fiboncci:")
fib
```



