---
title: "STAT 2200: Problem Set 1"
author: "Jack Ambery"
date: "Due: Thursday, 2/1 at the beginning of class"
fontsize: 12pt
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- set global options -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\vspace{0.1in}

```{=tex}
\begin{itemize}
  \item You may discuss this assignment with other students in the class, but 
    you may not sit down and type it up with them or show them your code. You 
    also may not discuss the assignment with anyone who isn't in our class, nor
    may you look up anything online.
  \item You must type up your homework using R Markdown. I want to see all of 
    your code and output, and any answers you provide that aren't code must be 
    typed above the respective \texttt{R} code chunk.
  \item Make sure none of your code runs off the page, otherwise you will lose 
    points.
  \item You must print and turn in a PDF of your homework. I won't accept 
    anything else (e.g., a Word document). All pages must be stapled together.
  \item This assignment is worth 150 points.
\end{itemize}
```
<!-- part 1 -->

\section{Part 1}

Type each expression below and print the resulting value. Do not store
anything.

\vspace{0.1in}

\noindent 1. $\quad \displaystyle 1 - 4 + 2 - 3 + 5 - 7 + 11$

```{r}
1 - 4 + 2 - 3 + 5 - 7 + 11
```

\noindent 2. $\quad \displaystyle \frac{4}{9}(96-31)$

```{r}
(4 / 9) * (96 - 31)
```

\noindent 3. $\quad \displaystyle \frac{21,000}{1-(1+0.035)^{-36}}$

```{r}
(21000) / (1 - ((1 + 0.035) ^ -36))
```

\noindent 4.
$\quad \displaystyle 2,500 \bigg(1+\frac{0.043}{12}\Bigg)^{12 \times 25}$

```{r}
(2500) * ((1 + (0.043 / 12)) ^ (12 * 25))
```

\noindent 5.
$\quad \displaystyle -\frac{1}{3} + \ln(24) + e^{3}\sqrt{\frac{3}{5} + 3 \times 5^{-2}}$

```{r}
(-1/3) + (log(24)) + ((exp(3)) * (sqrt((3 / 5) + 3 * (5 ^ -2))))
```

\newpage

<!-- part 2 -->

\section{Part 2}

\noindent 1. Create and store a vector of values starting at 5 and
ending at 73, with the values increasing in increments of 4.

```{r}
vec1 <- seq(from = 5, to = 73, by = 4)
```

\noindent 2. Add 2 to all of the values in the vector from problem 1 and
store the resulting vector. Do this by referring to the vector you made
in problem 1. Do not print anything.

```{r}
vec2 <- vec1 + 2
```

\noindent 3. Remove the seventh value from the vector in problem 2 and
store the new vector. Print the result.

```{r}
vec3 <- vec2[-7]
vec3
```

\noindent 4. Add the values 25 and 31 to the end of the vector in
problem 3 and store the new vector. Print the result.

```{r}
vec4 <- c(vec3, 25, 31)
vec4
```

\noindent 5. Suppose the fourth value in the vector from problem 4 was
supposed to be a 22. Make the appropriate change in the vector and print
the resulting vector.

```{r}
vec4[4] <- 22
vec4
```

\noindent 6. Sort the vector from problem 5 so that all of the values
are in increasing order. Then store the new vector. Do not print
anything here.

```{r}
vec5 <- sort(vec4)
```

\noindent 7. Using an \texttt{R} function, find the length of the vector
in problem 6.

```{r}
length(vec5)
```

\noindent 8. Create a new vector by repeating the number zero 40 times.
Print the result. Do not store this vector.

```{r}
rep(0, times = 40)
```

\noindent 9. Recreate the following vector using the \texttt{rep()}
function: 5, 5, 5, 1, 1, 1, 4, 4, 4, 2, 2, 2, 1, 1, 1. Also store it.

```{r}
vec6 <- c(rep(5, times = 3), rep(1, times = 3), rep(4, times = 3), 
          rep(2, times = 3), rep(1, times = 3))
```

\noindent 10. Combine the vectors from problems 5 and 9, and then print
the result. You do not need to store anything here.

```{r}
c(vec4, vec6)
```

\vspace{0.25in}

<!-- part 3 -->

\section{Part 3}

\noindent 1. The following are the final high school GPAs of a group of
college students: 3.63, 3.81, 3.96, 4.00, 3.12, 2.86, 3.77, 2.53, 3.26,
and 2.29. Create a vector that stores these values in the order
specified. Choose a variable name that's descriptive but relatively
short.

```{r}
hsGPAs <- c(3.63, 3.81, 3.96, 4.00, 3.12, 2.86, 3.77, 2.53, 3.26, 2.29)
```

\noindent 2. The following are the SAT quantitative scores for the same
group of students: 640, 660, 790, 770, 580, 560, 670, 440, 650, and 470,
respectively. Create another vector that stores these SAT scores in the
order provided.

```{r}
SATScores <- c(640, 660, 790, 770, 580, 560, 670, 440, 650, 470)
```

\noindent 3. Suppose I made a mistake when recording the SAT
quantitative score for the fifth student, whose actually score was a
680. Make this change using the vector you created in problem 2 and then
print the resulting vector.

```{r}
SATScores[5] <- 680
SATScores
```

\vspace{0.2in}

<!-- part 4 -->

\section{Part 4}

According to the International Monetary Fund (IMF), the countries with
the ten highest gross domestic products (GDPs) in 2018, from highest to
lowest, were the United States, China, Japan, Germany, the United
Kingdom, France, India, Italy, Brazil, and Canada. Their GDPs (in
millions of US dollars) were 20,513,000, 13,457,267, 5,070,626,
4,029,140, 2,808,899, 2,794,696, 2,689,992, 2,086,911, 1,909,386, and
1,733,706, respectively. \vspace{0.1in}

\noindent 1. Create a variable that consists of the GDPs specified above
and then print it. Use an appropriate name when storing, and be careful
with commas!

```{r}
gdps = c(20513000, 13457267, 5070626, 4029140, 2808899, 2794696, 2689992, 2086911, 
1909386, 1733706)
gdps
```

\noindent 2. Create a new variable that measures the GDPs in trillions
of US dollars by converting the GDPs you stored in problem 1. You must
use \texttt{R} to convert the values for you (recall: \texttt{R} is a
vectorized language). Then print the resulting vector.

```{r}
gdpsTrill = gdps / 1000000000
gdpsTrill
```

\noindent 3. What were Italy's and Brazil's GDPs in 2018? Use one of the
vectors you already created and subset it to print these values. You
must do this in one line of code! Then type your answers (in a full
sentence) above the \texttt{R} code chunk to practice typing text
outside of \texttt{R} code chunks.

The GDP's of Italy and Brazil in 2018 were \$2,086,911 and \$1,909,386,
respectively.

```{r}
gdps[8:9]
```

\noindent 4. What was the mean GDP (in 2018) of the ten countries listed
above? Calculate this value and then type a full sentence that answers
the question. This is meant to give you more practice typing text
outside of \texttt{R} code chunks.

The mean GDP of the 10 largest countries in 2018 was \$5,709,362.

```{r}
mean(gdps)
```

\noindent 5. What was the total (combined) GDP (in 2018) of the ten
countries listed above? Calculate this value and then type a full
sentence that answers the question. This is meant to give you even more
practice typing text outside of \texttt{R} code chunks.

The total combined GDPs of the 10 largest countries in 2018 (in
millions) was \$57,093,623.

```{r}
sum(gdps)
```

<!-- part 5 -->

\section{Part 5}

\noindent 1. Create and store a vector that consists of the following
movie lengths (in minutes): 94, 109, 110, 123, 125, 108, 92, 106, 84,
119, 110, and 140.

```{r}
durations <- c(94, 109, 110, 123, 125, 108, 92, 106, 84, 119, 110, 140)
```

\noindent 2. How many movie lengths are in the vector from problem 1?
Use an appropriate \texttt{R} function to determine this, and then type
your answer above your code (outside of the \texttt{R} code chunk).

The given vector has the lengths of 12 movies.

```{r}
length(durations)
```

\noindent 3. For each movie, determine if the movie is at least 90
minutes long. Your result should be a vector of logicals.

```{r}
durations >= 90
```

\noindent 4. Determine if any movies are over two hours long. Your
result should involve a logical.

Three movies have a duration that is over 2 hours.

```{r}
sum(durations > 120)
```

\noindent 5. Determine if all of the movies are shorter than 90 minutes.
Your result should involve a logical.

```{r}
sum(durations < 90) == length(durations)
```

\noindent 6. Print the movie lengths that are between 90 and 120
minutes, inclusive.

```{r}
durations[durations <= 120 & durations >= 90]
```

\noindent 7. Print the indices of the movie lengths that are between 95
and 115 minutes, inclusive.

```{r}
which(durations <= 115 & durations >= 95)
```

\noindent 8. Print the movie lengths that are at least 90 minutes but
not equal to 110 minutes.

```{r}
durations[durations >= 90 & durations != 110]
```

\noindent 9. Print the movie lengths that are either under 90 minutes or
over 120 minutes.

```{r}
durations[durations < 90 | durations > 120]
```

\noindent 10. Print the indices of the movie lengths that are exactly
either 123 or 110 minutes.

```{r}
which(durations == 123 | durations == 110)
```
